Multi-level modeling with Openflexo/\FML

### 2022-03-25 6-modelpresentation.tex
fix metamodel: cardinality of creator (between Actor and TaskType) was * and should be 1
format all requirements with new formatting

\begin{figure*}
 \centering
    % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel-1.5.pdf}
     \caption{Process management base metamodel (\texttt{Gateway}'s associations with \texttt{TaskType} are represented with attributes)}
    \label{fig:BaseMetamodel}
\end{figure*}

In this section, we present the base metamodel presented in
Fig.~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case,
whose partial description was provided in the challenge description. The
requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are
straightforwardly implemented by instantiating the \emph{XSure model}, as an
instance of base metamodel (the left side of
Fig.~\ref{fig:MultilevelArchitecture}).

@@

\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and
references a collection of \texttt{TaskType} through the composition relation
\texttt{taskTypes} with (0..*) cardinality (\textbf{P1}). A \texttt{TaskType}
is embedded in a \texttt{ProcessType} and inherits from its context. To
illustrate this in \textit{XSure} insurance domain use case, \textit{XSure
model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and
\inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of
\texttt{TaskType}.

\texttt{ProcessType} also references a collection of gateways, reified with the
\texttt{Gateway} concept hierarchy. \texttt{Gateway} is a specialization of
\texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit},
\texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\textbf{P2}).
Depending on its type and following its underlying operational semantics, a
gateway defines one or more inputs and one or more outputs. A
\texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType}
and a collection of final \texttt{TaskType} with both roles
\texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType}
(cardinality 0..*) (\textbf{P3}). \texttt{TaskType} exposes a creator role as a
reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same
conceptual level.

@@

The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the
\texttt{allowedActorTypes} relation to \texttt{ActorType} defined in
\texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors}
relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..*
cardinality). The same modeling pattern applies to \texttt{ArtefactType} also
extending the concept \texttt{Type}, and both relations
\texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in
\texttt{TaskType} (\textbf{P7}). \texttt{TaskType} additionally exposes an
\texttt{expectedDuration} role (expressed in number of days), satisfying
\textbf{P8}.

The \texttt{Actor} concept defines a boolean attribute called
\texttt{isSenior}, while \texttt{TaskType} defines an additional
\texttt{isCritical} boolean attribute, indicating that some instances are
flagged as critical and must be performed by senior actors. To fulfill
\textbf{P9} requirement, a supplementary constraint is required for
\texttt{TaskType} and is captured through the following invariant expressed in
the \FML language:
\begin{lstlisting}
forEach (actor : allowedActors) {
    assert !isCritical | actor.isSenior
}
\end{lstlisting}

@@

\texttt{Process} represents an enacted \texttt{ProcessType}, as defined in
previous subsection (\textbf{P10}). \FML defines behavioral features called
behavior. \texttt{ProcessType} defines the behavior
\texttt{newProcess(String)}, taking the name of the process to enact as
argument:

@@

This scheme relies on \FML dynamic binding mechanism to delegate to task types
the responsibility of instances creation. An instance of \texttt{Process}
references its unique type \texttt{ProcessType} by the specialized
\texttt{/type} role. Each instance of \texttt{TaskType} is ontologically
instantiated with a \texttt{Task} (\textbf{P11}), using the same factory
pattern where \texttt{TaskType} has the responsibility to manage the
ontological instantiation. A \texttt{Task} references its unique
\texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date}
basic role (\textbf{P12}).

The used and produced artifacts are implemented by roles
\texttt{usedArtifacts}, \texttt{producedArtifacts} and
\texttt{performingActors} defined in \texttt{Task} (\textbf{P13}). An instance
of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of
\texttt{ArtifactType} through the specialized \texttt{/types} role
(\textbf{P14} and \textbf{P16}). Likewise, \texttt{Actor} specializes
\texttt{MultiInstance} and references a set of \texttt{ActorType} through the
specialized \texttt{/types} relation (\textbf{P15}).

@@

Authorization for an actor to perform a task (\textbf{P17}) is captured either
by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes}
defined in \texttt{TaskType}. This mechanism is completed by the behaviors
\texttt{isAuthorizedActor}, \texttt{isValidActor} and \texttt{isValidActorType}
defined in \texttt{TaskType}:

@@

Actor types specialization is captured by the \texttt{parentActorTypes}
relation defined in \texttt{ActorType} (\textbf{P18}). This is completed by
both the definition of the \texttt{hasActorType(ActorType)} behavior in
\texttt{Actor} and the recursive behavior \texttt{isOrSpecializes(ActorType)}
in \texttt{ActorType}:

@@

Each concept inherits from \texttt{ModelingElement}, which defines a
\texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies
\textbf{P19} requirement.

@@

\inst{Requirements analysis} is an instance of \texttt{SETaskType},
\inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement
specifications} is an instance of \texttt{SEArtifactType}. The value of the
property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is
\texttt{\{}\inst{Requirement specifications}\texttt{\}} and is
\texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes}
(\textbf{S1}). Similarly, \inst{Test case design} is an instance of
\texttt{SETaskType}, with a value of \texttt{true} for its property
\texttt{isCritical}, and bound to \inst{Analyst} by the
\texttt{allowedActorTypes} relation. \inst{Test case design} produces
\inst{Test cases} instance of \texttt{SETaskType} (\textbf{S2}). The latter is
used in \inst{Test design review} (\texttt{producedArtifactTypes} relation)
(\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).

The capture of \textbf{S3} requirement is performed while defining
\inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a
relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
cardinality.  The \texttt{newTask(String)} behavior overrides the generic
implementation by instantiating a \texttt{CodingTask} concept, specializing
\texttt{Task}, as shown in the following excerpt of \FML code:

@@

The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a
relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
cardinality (\textbf{S4}).

\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or
more programming languages but produces code which is expressed in one language
only. This requirement is captured through the redefinition of the behavior
\texttt{declaresRequiredProducedArtifacts} where programming language should
also match.  \textbf{S6} is guaranteed though the \texttt{languages} relation
defined in \texttt{CodeArtifact} and the following invariant declared in
\texttt{CodeArtifact}:
\begin{lstlisting}
forEach (artifactType : types) {
  assert !(artifactType instanceof CodeArtifactType) | artifactType.doesImplement(languages)
}
\end{lstlisting}

\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:

@@

The requirement \textbf{S8} is simply expressed by the definition of the
\inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of
\texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType},
as shown in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.

A critical task must additionally produce artifacts that must be associated
with a validation task. This is modeled with a supplementary relation
\texttt{validatedWith} defined in \texttt{Artifact} and referencing
\texttt{Artifact} validating it (\textbf{S9}).

All software engineering artifacts defined in the context of Acme Software
Engineering Process are  instances of \texttt{SEArtifact}. This concept defines
two attributes: \texttt{responsible} (an \texttt{Actor} instance), and
\texttt{versionNumber} (an integer). It thus fulfills \textbf{S10}. \inst{Bob
Brown} is declared as an instance of \texttt{Actor}, and references
\inst{Analyst} and \inst{Tester} (\texttt{ActorType} instances). He is also
referenced by all instances of \texttt{SETaskType} as the creator for related
task types (\textbf{S11}).

For \textbf{S12}, we assume that all tasks may define an expected duration,
which might be checked during process execution. This is modeled by the
\texttt{expectedDuration} attribute in \texttt{TaskType}. Alternatively, this
could be modeled through the definition of a \texttt{SETesting} concept, as a
specialization of \texttt{SETaskType}, and the instantiation of \inst{Testing}
as an instance of \texttt{SETesting}. The business logic expressed by
\textbf{S12} requirement should then be redefined in \texttt{SETesting}.

\textbf{S13} requirement has been previously partially fulfilled. This must be
completed with an association between an artifact produced and the task that
validates it. This is modeled by the reference to \inst{Test Cases} as a
produced artifact type in \inst{Test Case design} and the reference to
\inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in
Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.

@@

The Base metamodel is completed with some behaviors implementing an execution
semantics for the executed processes. All tasks --- instantiated from
\texttt{TaskType} for a given enacted process --- manage a status. This status
is either \texttt{Not startable} (when not assigned to a performing actor or
when required input artifacts are not available), \texttt{Startable},
\texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or
\texttt{Completed}. A \texttt{Task} also manages a set of performing actors, a
begin and end date, some used and produced artifacts. The \texttt{Gateway}
business logic has also been implemented through the implementation of an
abstract \texttt{execute(Process)} behavior. The management of artifacts whose
semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment}
(\textbf{P13}) is also implemented.

>>>>

\begin{figure*}
 \centering
    % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
    %\includegraphics[width=1.0 \textwidth]{Figures/Metamodel-1.5.pdf}
    %fix creator cardinality (between Actor and TaskType)
    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel-17.pdf}
     \caption{Process management base metamodel (\texttt{Gateway}'s associations with \texttt{TaskType} are represented with attributes)}
    \label{fig:BaseMetamodel}
\end{figure*}

In this section, we present the base metamodel presented in
Fig.~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case,
whose partial description was provided in the challenge description. The
requirements \reqp{1} to \reqp{19} of XSure insurance domain are
straightforwardly implemented by instantiating the \emph{XSure model}, as an
instance of base metamodel (the left side of
Fig.~\ref{fig:MultilevelArchitecture}).

@@

\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and
references a collection of \texttt{TaskType} through the composition relation
\texttt{taskTypes} with (0..*) cardinality (\reqp{1}). A \texttt{TaskType}
is embedded in a \texttt{ProcessType} and inherits from its context. To
illustrate this in \textit{XSure} insurance domain use case, \textit{XSure
model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and
\inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of
\texttt{TaskType}.

\texttt{ProcessType} also references a collection of gateways, reified with the
\texttt{Gateway} concept hierarchy. \texttt{Gateway} is a specialization of
\texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit},
\texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\reqp{2}).
Depending on its type and following its underlying operational semantics, a
gateway defines one or more inputs and one or more outputs. A
\texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType}
and a collection of final \texttt{TaskType} with both roles
\texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType}
(cardinality 0..*) (\reqp{3}). \texttt{TaskType} exposes a creator role as a
reference to an \texttt{Actor} concept (\reqp{4}), which is at the same
conceptual level.

@@

The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the
\texttt{allowedActorTypes} relation to \texttt{ActorType} defined in
\texttt{TaskType} (with 0..* cardinality) captures \reqp{5} requirement.
Requirement \reqp{6} is symmetrically satisfied with \texttt{allowedActors}
relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..*
cardinality). The same modeling pattern applies to \texttt{ArtefactType} also
extending the concept \texttt{Type}, and both relations
\texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in
\texttt{TaskType} (\reqp{7}). \texttt{TaskType} additionally exposes an
\texttt{expectedDuration} role (expressed in number of days), satisfying
\reqp{8}.

The \texttt{Actor} concept defines a boolean attribute called
\texttt{isSenior}, while \texttt{TaskType} defines an additional
\texttt{isCritical} boolean attribute, indicating that some instances are
flagged as critical and must be performed by senior actors. To fulfill
\reqp{9} requirement, a supplementary constraint is required for
\texttt{TaskType} and is captured through the following invariant expressed in
the \FML language:
\begin{lstlisting}
forEach (actor : allowedActors) {
    assert !isCritical | actor.isSenior
}
\end{lstlisting}

@@

\texttt{Process} represents an enacted \texttt{ProcessType}, as defined in
previous subsection (\reqp{10}). \FML defines behavioral features called
behavior. \texttt{ProcessType} defines the behavior
\texttt{newProcess(String)}, taking the name of the process to enact as
argument:

@@

This scheme relies on \FML dynamic binding mechanism to delegate to task types
the responsibility of instances creation. An instance of \texttt{Process}
references its unique type \texttt{ProcessType} by the specialized
\texttt{/type} role. Each instance of \texttt{TaskType} is ontologically
instantiated with a \texttt{Task} (\reqp{11}), using the same factory
pattern where \texttt{TaskType} has the responsibility to manage the
ontological instantiation. A \texttt{Task} references its unique
\texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date}
basic role (\reqp{12}).

The used and produced artifacts are implemented by roles
\texttt{usedArtifacts}, \texttt{producedArtifacts} and
\texttt{performingActors} defined in \texttt{Task} (\reqp{13}). An instance
of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of
\texttt{ArtifactType} through the specialized \texttt{/types} role
(\reqp{14} and \reqp{16}). Likewise, \texttt{Actor} specializes
\texttt{MultiInstance} and references a set of \texttt{ActorType} through the
specialized \texttt{/types} relation (\reqp{15}).

@@

Authorization for an actor to perform a task (\reqp{17}) is captured either
by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes}
defined in \texttt{TaskType}. This mechanism is completed by the behaviors
\texttt{isAuthorizedActor}, \texttt{isValidActor} and \texttt{isValidActorType}
defined in \texttt{TaskType}:

@@

Actor types specialization is captured by the \texttt{parentActorTypes}
relation defined in \texttt{ActorType} (\reqp{18}). This is completed by
both the definition of the \texttt{hasActorType(ActorType)} behavior in
\texttt{Actor} and the recursive behavior \texttt{isOrSpecializes(ActorType)}
in \texttt{ActorType}:

@@

Each concept inherits from \texttt{ModelingElement}, which defines a
\texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies
\reqp{19} requirement.

@@

\inst{Requirements analysis} is an instance of \texttt{SETaskType},
\inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement
specifications} is an instance of \texttt{SEArtifactType}. The value of the
property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is
\texttt{\{}\inst{Requirement specifications}\texttt{\}} and is
\texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes}
(\reqs{1}). Similarly, \inst{Test case design} is an instance of
\texttt{SETaskType}, with a value of \texttt{true} for its property
\texttt{isCritical}, and bound to \inst{Analyst} by the
\texttt{allowedActorTypes} relation. \inst{Test case design} produces
\inst{Test cases} instance of \texttt{SETaskType} (\reqs{2}). The latter is
used in \inst{Test design review} (\texttt{producedArtifactTypes} relation)
(\reqs{1} and \reqs{13}, satisfied with \reqp{9}).

The capture of \reqs{3} requirement is performed while defining
\inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a
relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
cardinality.  The \texttt{newTask(String)} behavior overrides the generic
implementation by instantiating a \texttt{CodingTask} concept, specializing
\texttt{Task}, as shown in the following excerpt of \FML code:

@@

The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a
relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
cardinality (\reqs{4}).

\reqs{5} is more ambiguous as a task type \texttt{CodingTask} defines one or
more programming languages but produces code which is expressed in one language
only. This requirement is captured through the redefinition of the behavior
\texttt{declaresRequiredProducedArtifacts} where programming language should
also match.  \reqs{6} is guaranteed though the \texttt{languages} relation
defined in \texttt{CodeArtifact} and the following invariant declared in
\texttt{CodeArtifact}:
\begin{lstlisting}
forEach (artifactType : types) {
  assert !(artifactType instanceof CodeArtifactType) | artifactType.doesImplement(languages)
}
\end{lstlisting}

\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\reqs{7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:

@@

The requirement \reqs{8} is simply expressed by the definition of the
\inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of
\texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType},
as shown in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.

A critical task must additionally produce artifacts that must be associated
with a validation task. This is modeled with a supplementary relation
\texttt{validatedWith} defined in \texttt{Artifact} and referencing
\texttt{Artifact} validating it (\reqs{9}).

All software engineering artifacts defined in the context of Acme Software
Engineering Process are  instances of \texttt{SEArtifact}. This concept defines
two attributes: \texttt{responsible} (an \texttt{Actor} instance), and
\texttt{versionNumber} (an integer). It thus fulfills \reqs{10}. \inst{Bob
Brown} is declared as an instance of \texttt{Actor}, and references
\inst{Analyst} and \inst{Tester} (\texttt{ActorType} instances). He is also
referenced by all instances of \texttt{SETaskType} as the creator for related
task types (\reqs{11}).

For \reqs{12}, we assume that all tasks may define an expected duration,
which might be checked during process execution. This is modeled by the
\texttt{expectedDuration} attribute in \texttt{TaskType}. Alternatively, this
could be modeled through the definition of a \texttt{SETesting} concept, as a
specialization of \texttt{SETaskType}, and the instantiation of \inst{Testing}
as an instance of \texttt{SETesting}. The business logic expressed by
\reqs{12} requirement should then be redefined in \texttt{SETesting}.

\reqs{13} requirement has been previously partially fulfilled. This must be
completed with an association between an artifact produced and the task that
validates it. This is modeled by the reference to \inst{Test Cases} as a
produced artifact type in \inst{Test Case design} and the reference to
\inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in
Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.

@@

The Base metamodel is completed with some behaviors implementing an execution
semantics for the executed processes. All tasks --- instantiated from
\texttt{TaskType} for a given enacted process --- manage a status. This status
is either \texttt{Not startable} (when not assigned to a performing actor or
when required input artifacts are not available), \texttt{Startable},
\texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or
\texttt{Completed}. A \texttt{Task} also manages a set of performing actors, a
begin and end date, some used and produced artifacts. The \texttt{Gateway}
business logic has also been implemented through the implementation of an
abstract \texttt{execute(Process)} behavior. The management of artifacts whose
semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment}
(\reqp{13}) is also implemented.

### 2022-03-25 7-requirementssatisfaction.tex
format all requirements with new formatting

\begin{itemize}
    \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to
      fulfill or meet the requirement (for example for \textbf{P1}:
      \textit{process type} is conceptualized in a concept
    \texttt{ProcessType}).

    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in
      the sense of object-oriented modeling (for example for \textbf{P2}:
      \texttt{Gateway} defines an abstract behavior \texttt{execute()} and
    \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit}
  and \texttt{OrJoin} are defined as inherited concepts implementing specific
business logic).

    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation,
      or specific attributes were added to some concepts (for example for
      \textbf{P3}: a \textit{process type} has one \textit{initial task type}).

    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is
      satisfied by the instantiation of one or more concepts (for example for
      \textbf{S1}: \inst{Requirement analysis} is defined as an instance of
    \texttt{SETaskType}).

    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a
      requirement is obtained by the relation between an instance and its type
      definition instance (for example the notion of \textit{developer} of
      \textbf{S3} is implemented by both a concept in \textit{Acme metamodel}
    and an instance in \textit{Acme model}, see bottom left of
  Fig.~\ref{fig:AcmeFullArchitecture}).

    \item \emph{Behavioral modeling} (\textit{B.Mod.}): a requirement is
  satisfied by the definition of one or more behaviors, which may be combined
  with constraints checking and associated tools (for example \textbf{P17}).
  \end{itemize}

@@

\renewcommand{\arraystretch}{1.1}
\begin{table*}
 \centering
 \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
   \hline
    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
   \cline{1-7}
                 & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.   &  checking & \\
  \hline
    \textbf{P1}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \textbf{P2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
    \textbf{P3}  &            &            & \checkmark &            &            &            &            & \\
    \textbf{P4}  & \checkmark &            &            &            & \checkmark &            &            & \\
    \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \textbf{P6}  & \checkmark &            & \checkmark &            &  &            &            & \\
    \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \textbf{P8}  &            &            & \checkmark &            &            &            &            & \\
    \textbf{P9}  &            &            & \checkmark &            &            &            & \checkmark & \\
    \textbf{P10} & \checkmark &            &            &            & \checkmark &            &            & \\
    \textbf{P11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
    \textbf{P12} &            &            & \checkmark &            &            &            &            & \\
    \textbf{P13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
    \textbf{P14} & \checkmark &            &            &            & \checkmark &            &            & \\
    \textbf{P15} & \checkmark &            &            &            & \checkmark &            &            & \\
    \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
    \textbf{P17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
    \textbf{P18} & \checkmark           &            &            &            &            & \checkmark &            & \\
    \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
  \hline
\end{tabular}
     \caption{Requirements satisfaction for base metamodel}
    \label{tab:RequirementsSatisfactionBaseMetamodel}
\end{table*}

\begin{table*}
 \centering
  \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
   \hline
    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
   \cline{1-7}
                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      & checking   & \\
  \hline
    \textbf{S1}  &            &            &            & \checkmark &            &            &            & \\
    \textbf{S2}  &            &            &            & \checkmark &            &            &            & \\
    \textbf{S3}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
    \textbf{S4}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \\
    \textbf{S5}  &            & \checkmark &            & \checkmark &            & \checkmark &            & \\
    \textbf{S6}  &            & \checkmark &            & \checkmark &            & \checkmark & \checkmark & \\
    \textbf{S7}  &            &            &            & \checkmark &            & \checkmark &            & \\
    \textbf{S8}  &            &            &            & \checkmark &            &            &            & \\
    \textbf{S9}  &            &            &            &            &            & \checkmark & \checkmark & \\
    \textbf{S10} & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
    \textbf{S11} &            &            &            & \checkmark &            &            &            & \checkmark \\
    \textbf{S12} &            &            &            & \checkmark &            &            &            & \checkmark \\
    \textbf{S13} &            &            &            & \checkmark &            &            &            & \\
  \hline
\end{tabular}
     \caption{Requirements satisfaction for Acme software engineering process}
    \label{tab:RequirementsSatisfactionAcme}
\end{table*}

>>>>

\begin{itemize}
    \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to
      fulfill or meet the requirement (for example for \reqp{1}:
      \textit{process type} is conceptualized in a concept
    \texttt{ProcessType}).

    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in
      the sense of object-oriented modeling (for example for \reqp{2}:
      \texttt{Gateway} defines an abstract behavior \texttt{execute()} and
    \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit}
  and \texttt{OrJoin} are defined as inherited concepts implementing specific
business logic).

    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation,
      or specific attributes were added to some concepts (for example for
      \reqp{3}: a \textit{process type} has one \textit{initial task type}).

    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is
      satisfied by the instantiation of one or more concepts (for example for
      \reqs{1}: \inst{Requirement analysis} is defined as an instance of
    \texttt{SETaskType}).

    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a
      requirement is obtained by the relation between an instance and its type
      definition instance (for example the notion of \textit{developer} of
      \reqs{3} is implemented by both a concept in \textit{Acme metamodel}
    and an instance in \textit{Acme model}, see bottom left of
  Fig.~\ref{fig:AcmeFullArchitecture}).

    \item \emph{Behavioral modeling} (\textit{B.Mod.}): a requirement is
  satisfied by the definition of one or more behaviors, which may be combined
  with constraints checking and associated tools (for example \reqp{17}).
  \end{itemize}

@@

\renewcommand{\arraystretch}{1.1}
\begin{table*}
 \centering
 \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
   \hline
    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
   \cline{1-7}
                 & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.   &  checking & \\
  \hline
    \reqp{1}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \reqp{2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
    \reqp{3}  &            &            & \checkmark &            &            &            &            & \\
    \reqp{4}  & \checkmark &            &            &            & \checkmark &            &            & \\
    \reqp{5}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \reqp{6}  & \checkmark &            & \checkmark &            &  &            &            & \\
    \reqp{7}  & \checkmark &            & \checkmark &            &            &            &            & \\
    \reqp{8}  &            &            & \checkmark &            &            &            &            & \\
    \reqp{9}  &            &            & \checkmark &            &            &            & \checkmark & \\
    \reqp{10} & \checkmark &            &            &            & \checkmark &            &            & \\
    \reqp{11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
    \reqp{12} &            &            & \checkmark &            &            &            &            & \\
    \reqp{13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
    \reqp{14} & \checkmark &            &            &            & \checkmark &            &            & \\
    \reqp{15} & \checkmark &            &            &            & \checkmark &            &            & \\
    \reqp{16} & \checkmark &            &            &            & \checkmark &            &            & \\
    \reqp{17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
    \reqp{18} & \checkmark           &            &            &            &            & \checkmark &            & \\
    \reqp{19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
  \hline
\end{tabular}
     \caption{Requirements satisfaction for base metamodel}
    \label{tab:RequirementsSatisfactionBaseMetamodel}
\end{table*}

\begin{table*}
 \centering
  \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
   \hline
    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
   \cline{1-7}
                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      & checking   & \\
  \hline
    \reqs{1}  &            &            &            & \checkmark &            &            &            & \\
    \reqs{2}  &            &            &            & \checkmark &            &            &            & \\
    \reqs{3}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
    \reqs{4}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \\
    \reqs{5}  &            & \checkmark &            & \checkmark &            & \checkmark &            & \\
    \reqs{6}  &            & \checkmark &            & \checkmark &            & \checkmark & \checkmark & \\
    \reqs{7}  &            &            &            & \checkmark &            & \checkmark &            & \\
    \reqs{8}  &            &            &            & \checkmark &            &            &            & \\
    \reqs{9}  &            &            &            &            &            & \checkmark & \checkmark & \\
    \reqs{10} & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
    \reqs{11} &            &            &            & \checkmark &            &            &            & \checkmark \\
    \reqs{12} &            &            &            & \checkmark &            &            &            & \checkmark \\
    \reqs{13} &            &            &            & \checkmark &            &            &            & \\
  \hline
\end{tabular}
     \caption{Requirements satisfaction for Acme software engineering process}
    \label{tab:RequirementsSatisfactionAcme}
\end{table*}
